---
title: "Project 2"
author: "Andrew Eeckman | Beitong Li | Martin Marquez | Sriman Manyam"
date: "03/04/2021"
output:
  html_document:
    code_folding: "hide"
    theme: "spacelab"
---

```{r,include=FALSE,results='hide'}
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggthemes)
library(patchwork)
library(gganimate)
```


# Abstract

This report details efforts to analyze how rising levels of government stringency in response to COVID-19 are associated with its spread throughout a given country. Our hypothesis is that those with stricter responses typically experience fewer new cases than those with more lenient responses. Using data gathered by the World Health Organization (WHO) on COVID-19, we are able to address this question using the most reliable and up to date data available on COVID-19. (We should describe the problem, key variables, and possible hypotheses in a couple sentences here.)


***

# Introduction


The COVID-19's impact on our world has been significant having in resulted in large losses of life throughout several countries, massive challenges to public health infrastructures, a wide array of negative economic set-backs, and much more. The worldwide challenge, since the rise of COVID-19, has been how to quickly address and remedy this pandemic. Using data from the WHO and Oxford University on COVID-19's spread and the policies used to counteract it, this report will seek to analyze how the stringency of government response is associated with the spread of COVID-19. Written formally, our primary question of interest is the following:

Are rising levels of government stringency in response to COVID-19 associated with a decrease in its spread throughout a given country?

Our hypothesis to this above question is that stricter responses are associated with a decreasing values of new cases in a given country. 


***

# Background

The data used in this reports comes primarily from the World Health Organization's Covid-19 global dataset, which is updated daily.  Because it is updated daily, our data will include the data collected as of March 4th, 2020. The WHO Covid-19 global dataset serves as a central repository for worldwide Covid-19 data, and provides data on Covid-19 cases from January 3rd 2020 to the present day. Specifically, this dataset includes a daily record of each country/region's new cases, cumulative cases up to that date, new deaths, and total deaths in that country/region up to that date. In other words, it provides a simplified record of covid-19  cases and deaths, organized by date and by country/region. A clarification here would be, that some regions are counted separately from their ruling countries. Most of these cases are: colonized islands, autonomous religious regions, or disputed lands, for which the individualization of these regions is helpful as the covid-19 response may be different in those regions, even if they are under the rule of a country already represented. Therefore, this dataset covers 426 days and 237 countries/regions, when we multiply them we get the 100,962 observations we have in our dataset. The data encompasses all individuals who have contracted and/or perished from covid-19, and is collected directly by each country and then reported to the WHO. This presents room for potential inaccuracies, as different countries may measure Covid-19 differently, and some countries may report inaccurate numbers, not measure cases at all, or deliberately withhold numbers altogether. Generally, the expected issue will lead to potential case underreporting, and that is helpful to keep oin mind. However, considering the scale of the data encompasses nearly the entire human population, and is dependent on respective governments, this data is about as accurate as anyone could hope to feasibly acquire for these purposes, and the large sample size should help to counteract. Moreover, cases where countries are not recording or reporting their numbers, will not be included in this analysis. The WHO global dataset can be found [here] (https://covid19.who.int/WHO-COVID-19-global-data.csv)

A secondary dataset is used in this report, and is merged with the aforementioned dataset. This dataset can be found in ourworldindata.org, which is an established scientific publication, and originates from research conducted by Oxford University. This dataset is likewise updated daily, and began measuring observations on Jan. 1st 2020. The dataset is grouped by country and date, and indicates each observations respective "containment index", which is a score assigned to each country for that date, which grades in a number between 0-100, how well that country's government responded to Covid-19. The measure is based off an additive score of nine indicators, such as school closures, travel bands, etc..., and will be used in its intended context of comparing with one another, rather than an objective metric for appropriate response.  More information for this data can be found [here] (https://ourworldindata.org/grapher/covid-containment-and-health-index?tab=table&stackMode=absolute&time=earliest..latest&region=World)


***


```{r,message=FALSE,results='hide',warning=FALSE,include=FALSE}

# SET UP

covid <- read_csv("../Data/https://covid19.who.int/WHO-COVID-19-global-data.csv")
c_h_index <- read_csv("../Data/covid-containment-and-health-index.csv")
pop_index <- read_csv("../Data/population.csv")

colnames(covid)[which(names(covid) == "Date_reported")] <- "Date"
colnames(covid)[which(names(covid) == "Country_code")] <- "Code"

colnames(c_h_index)[which(names(c_h_index) == "Entity")] <- "Country"

covid <- covid[!(covid$Date == "2021-03-03" | covid$Date == "2021-03-02"),]
covid$Date_of_Restrictions <- covid$Date - 28

covid_index <- merge(covid, c_h_index, by.x = c("Date_of_Restrictions", "Country"), by.y = c("Date", "Country"))

pop <- pop_index[c("Country Name","2019")]
pop <- pop[!is.na(pop["2019"]),]

covid_index_pop = merge(covid_index, pop, by.x = c("Country"), by.y = c("Country Name"))

colnames(covid_index_pop)[which(names(covid_index_pop) == "2019")] <- "Population"

covid_index_pop$new_cases_div_by_pop <- covid_index_pop$New_cases / covid_index_pop$Population
```

***

# Basic Summarization

In this section, we focus on visualize the top 3 countries with the most accumulated Covid new cases since the pandemic. The purpose is to address the severity of the spreading speed and emphasize on the major Covid-impact areas. Before we jump into the major analysis, we would like to provide readers with the basic summarization, including number of new cases, number of death, Infection Fatality Ratio. 

Up to Feburary 28th, 2021 Worldwide Level:

- Accumulated New Covid Cases `sum(covid$New_cases)`: 114853685
  
- Accumulated Death Covid Cases `sum(covid$New_deaths)`: 2554694
  
- Infection Fatality Ratio (IFR, in %)： 2.224%

- Infection Fatality Ratio (IFR, in %) = Number of deaths from disease / Number of infected individuals * 100

```{r,include=FALSE,results='hide'}
# total cases worldwide to date
total.cc = sum(covid$New_cases)
a <- covid %>% group_by(Country) %>% 
 summarise(sum_newcase = sum(New_cases), max_newcase_perday_bycountry = max(New_cases)) %>% 
 arrange(desc(sum_newcase))
# total deaths worldwide to date
total.cd = sum(covid$New_deaths)
b <- covid %>% 
 group_by(Country) %>% 
 summarise(sum_newdeath = sum(New_deaths), max_newdeath_perday_bycountry = max(New_deaths)) %>% 
 arrange(desc(sum_newdeath))

# I'm not sure how to include the above section in, it gives us information that the section below I think best elaborates on.
```

In order to address to what degree a government's response may associate with Covid-19's spread, it is prudent to first demonstrate that there are measurable differences in Covid-19's spread, by country. These differences in Covid-19's spread are generally known about by the public, and were the biggest source of inspiration for this report.

```{r,message=FALSE,results='hide'}
#head(covid)

plot1 <- ggplot(data=covid, mapping = aes(x=Date, y=New_cases)) + 
  geom_line(color = "steelblue2", size = 0.1) +
  labs(title = "Worldwide Total New Covid Cases Per day",
       subtitle = "How is Covid spreading around the world?",
       x = "Date Reported (Daily)",
       y = "# of New Covid Cases") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text())

plot2 <- ggplot(data=covid, mapping = aes(x=Date, y=New_deaths)) + 
  geom_line(color = "sienna2", size = 0.1) +
  labs(title = "Worldwide Total Covid Death Cases Per day",
       subtitle = "How many people died from Covid?",
       x = "Date Reported (Daily)",
       y = "# of New Covid Deaths") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) 

plot1 / plot2 + plot_layout(ncol = 1, heights=c(4,4))
```

The plots above us, give us an idea of how Covid-19 deaths may relate to Covid-19's spread. Intuitively, we may expect that higher amounts of Covid-19 cases would lead to higher amounts of Covid-19 deaths. However, looking at Worldwide Covid Death Cases per Day, we can see a sharp spike around April and May 2020, where Worldwide New Covid Cases per Day doesn't seem to indicate a particular spike in cases. How and why can this be? We will take a look at the country of Italy below, which may be a significant source of the spike in deaths.

```{r,message=FALSE,results='hide'}

Italy <- covid[covid$Country == "Italy",]

plot_Italy_c <- Italy %>% ggplot(aes(x=Date, y=New_cases)) + 
  geom_line(color = "darkorange", size = 0.5)+
   labs(title = "Total Covid New Cases in Italy (Daily)",
       subtitle = "How covid cases increase in the Italy?",
       x = "Date Reported (Daily)",
       y = "# of New Cases") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) 

plot_Italy_d <- Italy %>% ggplot(aes(x=Date, y=New_deaths)) + 
  geom_line(color = "firebrick4", size = 0.5)+
   labs(title = "Total New Covid Deaths in Italy (Daily)",
       subtitle = "How covid deaths increase in the Italy?",
       x = "Date Reported (Daily)",
       y = "# of New Deaths",
       color = "Country") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) 


plot_Italy_c / plot_Italy_d + plot_layout(ncol = 1, heights=c(4,4))
```

The above plots show something interesting around the months of April and May. We can see a small increase in Covid-19 cases, led to a very large increase in related deaths. For a local comparison, we can see cases between Nov. 2020 and Jan. 2021 were between two and four times as high as they were during April/May, yet deaths soon following those case spikes about match the death totals in April/May. This suggests there are factors at work, that may be resulting in these kinds of variations.

For a bit of foreshadowing, we can show what the containment index was prior to the sharp spike in deaths in April/May. We can see the containment index was not too high, prior to the spike in cases. To what degree a less adequate response may associate with covid cases, we can't say definitively yet. This only serves as an interesting example to highlight the portions of the data that may prove significant in testing our question of interest.

```{r,message=FALSE,results='hide'}

Italy2 <- covid_index[covid_index$Country == "Italy",]

plot_Italy_ci <- Italy2 %>% ggplot(aes(x=Date, y=containment_index)) + 
  geom_line(color = "lightpink4", size = 0.5)+
   labs(title = "Containment Index score in Italy (Daily)",
       subtitle = "How the containment inbdex changes in the Italy?",
       x = "Date Reported (Daily)",
       y = "containment index (0-100)") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) 

plot_Italy_d <- Italy %>% ggplot(aes(x=Date, y=New_deaths)) + 
  geom_line(color = "firebrick4", size = 0.5)+
   labs(title = "Total New Covid Deaths in Italy (Daily)",
       subtitle = "How covid deaths increase in the Italy?",
       x = "Date Reported (Daily)",
       y = "# of New Deaths") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text())
  


plot_Italy_d / plot_Italy_ci + plot_layout(ncol = 1, heights=c(4,4))
```

Below, we can see daily Covid-19 reported cases in the 3 countries with the largest amount of cases. We can see spikes and valleys in the plots at different times. In the US, cases spike around Dec. 2020, where in India they might spike around Sept. 2020. This suggests case spikes are due more to localized reasons, and that each country roughly experiences a unique pattern of spread. An obvious question that arises, is to what degree the population size may explain the higher amount of cases. This is a factor to some degree, as the US and India are the third and second largest countries by population respectively, however India has a population roughly 4 times the size of the US population and slightly less total cases than the US. Moreover, China with a population also about 4 times larger than the US, is rank 85 in regards to total Covid-19 cases. Brazil currently has the third highest amount of Covid-19 cases, and is the sixth largest country by population. This means three countries larger in population, have fewer cases, in some cases far fewer cases. All this to suggest, that larger factors such than population size alone, such as government response, may be at play.

```{r,message=FALSE,results='hide',warning=FALSE}

US <- covid[covid$Country == c("United States of America","Brazil","India"),] 

plot_top3 <- US %>% ggplot(aes(x=Date, y=New_cases,color = Country)) + 
  geom_line(size = 0.5)+
   labs(title = "Total Covid New Cases in Top Three Countries (Daily)",
       subtitle = "How covid increases in Top Three Countries?",
       x = "Date Reported (Daily)",
       y = "# of New Cases") + 
  theme(axis.title = element_text()) + 
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) + 
  theme(plot.title = element_text(size=16))
  
plot_top3

```

The United States of America, Brazil, and India are the top 3 countries which had accumulated over 10,000,000 Covid diagnosis cases since the beginning of the pandemic. We calculated a total of 114853685 diagnosis Covid cases, and the top three countries contributed to 43.7% of the total world wide new Covid cases. This data could serve as a warning alert for WHO to focus on helping countries with the most cases and upgrade the system of Covid vrius detection and prevention.

*** 

# DESCRIPTIVE ANALYSIS

Based on our primary question of interest, the variables that we are interested in exploring from these two data sets includes `New_cases`, which tracks the number of new cases found in a country each day, and `containment_index`, which measures the stringency of said country's response to COVID-19. Since we are interested in analyzing these two variables together, we'll need to merge the two datasets they are found in. 

However, before we do so we have to recognize an important fact about how COVID-19 spreads and its potential relationship to government restrictions. COVID-19 infections that are symptomatic and thus the most contagious can last up to four weeks or more. People sick today might not know they are sick until two weeks after they contract the virus and they may infect others for up to several weeks later. This means that any restrictions or response made today is not likely to impact the number of new cases for several weeks. For the purpose of this report we will thus be comparing the number of new cases on any given day with the containment and health index from four weeks prior. This way we can avoid associating government responses with new cases they could not have yet potentially influenced.

We should also take into account another important factor in the spread of COVID-19, population. Countries with larger populations are likely to have more cases, and thus, more new cases per day than smaller countries following similar guidelines. Thus, we should standardize the number of new cases per day for each country by dividing said value by that countries population. Merging our newly created dataset with 2018 population data, we obtain a final dataset that contains not only the number of new cases per day for each country but also each country's population and containment index. Finally, we will divide the number of new cases per day for each country by that country's population to get the main variable used for our analysis, `news_cases_div_by_pop`. 


```{r}

covid$Date_of_Restrictions <- covid$Date - 28

covid_index <- merge(covid, c_h_index, by.x = c("Date_of_Restrictions", "Country"), by.y = c("Date", "Country"))

pop <- pop_index[c("Country Name","2019")]
pop <- pop[!is.na(pop["2019"]),]

covid_index_pop = merge(covid_index, pop, by.x = c("Country"), by.y = c("Country Name"))

colnames(covid_index_pop)[which(names(covid_index_pop) == "2019")] <- "Population"

covid_index_pop$new_cases_div_by_pop <- covid_index_pop$New_cases / covid_index_pop$Population

```

The variable `containment_index` has values that range from 0 to 100, with 0 meaning that no restrictions and policies have been put in place and with 100 meaning that the strictest restrictions and policies have been adopted. As of right now this variable is a continuous one. But for the purposes of our project, we want to be able to categorize a country's stringency as compared to the rest of its peers. Thus, we must first find a set of values we perceive to be representative of the many levels of stringency around the world and then we must transform the `containment_index` variable into a categorical one. 

Let us start by first visualizing the distribution of this variable starting at earliest date available to us given the WHO dataset (01/03/2020)

```{r}
# covid-19 01/03/2020

plot_01_03_2020 <- covid_index_pop %>% ggplot() + 
  geom_histogram(mapping = aes(x = containment_index), fill = "seagreen4", size = 0.5, binwidth = 1)+
   labs(title = "Containment Index Distribution starting on 01/03/2020",
       x = "Containment Index",
       y = "Frequency") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) +
  theme(plot.title = element_text(size=16))
plot_01_03_2020

```


As we can see from above, there is one value that stands out starkly from the rest, 0. Now this could be because many countries simply chose not to respond to COVID throughout the pandemic or it could be because of the dates we chose to start at. On 01/03/2020, COVID-19 was not yet considered a global threat and so it makes sense that for most countries zero restrictions and policies would be in place. To confirm this theory, lets take a look at how the distribution changes when only values after 01/30/2020 are considered (01/30/2020 is the day the WHO announced COVID-19 was of global concern).


```{r,include=FALSE,message=FALSE}
# covid-19 01/30/2020

covid_post_0130 <- with(covid_index_pop, covid_index_pop[Date > "2020-01-29",])
covid_post_0130

```

```{r}

plot_01_30_2020 <- covid_post_0130 %>% ggplot() + 
  geom_histogram(mapping = aes(x = containment_index), fill = "seagreen4", size = 0.5, binwidth = 1)+
   labs(title = "Containment Index Distribution starting on 01/30/2020",
       x = "Containment Index",
       y = "Frequency") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) +
  theme(plot.title = element_text(size=16))
plot_01_30_2020

```


The distribution above has not changed much at all and a containment index of 0 still stands out starkly from the rest. Perhaps the first theory mentioned above is correct and many governments have just not responded to COVID-19. However, we should keep in mind that on 01/30/2020, there were only roughly 6,102 cases worldwide with the majority being in China. It is possible that due to a lack of domestic cases, many countries had not yet started paying attention to COVID-19. For these reasons, let's take a look at an even later date, 03/11/2020, the day COVID-19 was declared a global pandemic by the WHO. The containment index distribution taken after from that date onward is as follows:

```{r,include=FALSE,message=FALSE}
#On March 11th, 2020, the WHO declared COVID-19 a global pandemic
# covid-19 03/11/2020

covid_post_0311 <- with(covid_index_pop, covid_index_pop[Date > "2020-03-10",])
covid_post_0311
```

```{r}

plot_03_11_2020 <- covid_post_0311 %>% ggplot() + 
  geom_histogram(mapping = aes(x = containment_index), fill = "seagreen4", size = 0.5, binwidth = 1)+
   labs(title = "Containment Index Distribution starting on 03/11/2020",
       x = "Containment Index",
       y = "Frequency") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) +
  theme(plot.title = element_text(size=16))
plot_03_11_2020
```


As we can see the number of 0 index values has fallen sharply by 03/11/2020. This means that following the announcement of COVID-19 as a global pandemic, most countries had started responding to it, imposing restrictions and policies designed to combat its spread. In fact, we can see in the below graph that the vast majority of countries before this date largely did nothing to address the virus. 

```{r,include=FALSE}

# covid-19 01/30/2020 - 3/10/2020

covid_post_01300310 <- with(covid_index_pop, covid_index_pop[Date > "2020-01-29" & Date < "2020-03-10",])
covid_post_01300310
```

```{r}
# Plot for ^^
plot_01_30_thru_03_11_2020 <- covid_post_01300310 %>% ggplot() + 
  geom_histogram(mapping = aes(x = containment_index), fill = "seagreen4", size = 0.5, binwidth = 1)+
   labs(title = "Containment Index Distribution 01/30/2020 thru 03/11/2020",
       x = "Containment Index",
       y = "Frequency") +
  theme_fivethirtyeight() + 
  theme(axis.title = element_text()) +
  theme(plot.title = element_text(size=16))
plot_01_30_thru_03_11_2020
```


With all that being said, using the containment index distribution sourced from dates after the announcement is more likely to be better representative of how governments actually responded to this pandemic (again, before this date, it technically was not a pandemic yet). Seeing as we want to split the data into relatively equal parts, percentiles are likely the best possible choice in this case. Listed below are the percentile values for the 20th, 40th, 60th, and 80th percentiles. 

```{r}
#Lenient, Lenient-Moderate, Moderate, Strict-Moderate, Strict

quantile(covid_post_0311$containment_index, c(0.2, 0.4, 0.6, 0.8))

```


Categorizing the containment index by these values we end up with 5 different levels and a new variable called `containment_index_leveled`. 

- For containment index values from 0 to 38.46, countries are labeled as being lenient (Lenient) in their responses. 

- For containment index values from 38.46 to 51.92, countries are labeled as being moderately lenient (Lenient_Mod) in their responses. 

- For containment index values from 51.92 to 61.54, countries are labeled as being moderate (Moderate) in their responses. 
- For containment index values from 61.54 to 69.23, countries are labeled as being moderately strict (Mod_Strict) in their responses. 

- For containment index values from 69.23 to 100, countries are labeled as being strict (Strict) in their responses. 

```{r,message=FALSE,results='hide',include=FALSE}
covid_index_pop$containment_index_leveled <- cut(covid_index_pop$containment_index,
                                                 breaks = c(-Inf, 38.46, 51.92, 61.54, 69.23 , Inf),
                                                 labels = c("Lenient", "Lenient_Mod", "Moderate", "Mod_Strict", "Strict"))


```

With `containment_index` transformed into a categorical variable, we can go on to analyze its relationship to new cases using a two way ANOVA model.

*** 

# Inferential Analysis

In order to assess if there is in fact a correlation between a country's COVID-19 response and the spread of COVID-19, we will construct a two way ANOVA model. ANOVA models are specifically designed to test the influence of a categorical variable, in this case the leveled containment index, on a continuous one, the number of new COVID-19 cases. While this makes ANOVA modeling quite useful, several assumptions need to be made before using these models.

The first, and most straight forward, of these assumptions is that the data points that are used are relevant in regards to the question being asked. The containment index levels we are using is derived from the previous quantiles listed above. We transformed containment index, a continuous variable, into a categorical one, separated the values into 5 distinct levels that correlate to a country's COVID-19 response (Lenient, Lenient-Moderate, Moderate, Moderate-Strict, and Strict). Since we are testing to see whether or not a country's COVID-19 is associated with how the virus spreads, it is clear that both `Country` and `containment_index_leveled` are relevant to the question being asked. The second assumption of this model is that the response variable is continuous. This assumption is definitely satisfied as we know the response variable, new cases of COVID-19 scaled by the countries population, is a continuous variable as it is calculated by dividing the number of new cases by the total population for each country. The third assumption is that this response variable follows a normal distribution. As shown previously, this is also likely satisifed. The fourth and final assumption is that the variance of the data in different groups is the same. This can be proved to be satisfied by the sensitivity analysis portion of the report. Our containment level variable is based off of the spread of data found in the above histogram, as mentioned previously, and is separated accordingly so that the variance of the data in each level (Lenient to Strict) is the same. 

Finally, we can begin constructing the following ANOVA model. The null hypothesis for this model, in layman’s terms, is that containment index level, regardless of what it is, is not correlated with a change in the number of new COVID-19 cases. The alternative hypothesis, on the other hand, states that at least one containment index level is indeed correlated with changes to a country's number of new COVID-19 cases.
  
```{r}
two_way_anova <- aov(new_cases_div_by_pop~containment_index_leveled+Country, data = covid_index_pop)
summary(two_way_anova)
```

At a 95% significance level, our table provides an F-value of 728.06 for `containment_index_leveled` and an F-value of 95.19 for `Country`. These two variables both have p-values of < 2e-16 as returned by the above ANOVA model. This low p-value indicates that we can reject the null hypothesis at a 95% significance level, allowing us to state that the `containment_index_leveled` is correlated with the changes in the number of new COVID-19 cases for a particular country.

At this point since we have now confirmed that there are statistical differences between containment levels, it would benefit us to quantify these differences using a Tukey's test. The Tukey's test, similar to the ANOVA model, makes multiple assumptions about the tested data. First, the Tukey's test assumes that the observations are independent within and among differing groups. As stated previously, we know this is the case because of how we constructed the levels, using quantile based sectioning. The next assumption is that for each mean in the tested data, the groups associated are normally distributed. Since we know that the mean COVID spread rate approximately follows a normal distribution, thereby likely satisfying this assumption. The final assumption is that there is homogeneity of variance across the groups associated with each mean in the test. This was shown above when we tested each containment level. 

Now that we've satisfied all assumptions from the Tukey's test, we can construct it.

```{r}
TukeyHSD(two_way_anova, "containment_index_leveled")
```

At a 95% confidence level we see that there are statistically significant differences between the following containment levels in terms of their associated changes in the number of new COVID-19 cases per day. 

- Lenient with Lenient-Moderate, Moderate, Moderate-Strict, and Strict
- Lenient-Moderate with Moderate, Moderate-Strict, and Strict
- Moderate with Moderate-Strict and Strict
- Moderate-Strict with Strict

Something curious to note is that when moving from more lenient containment levels to more stricter ones, with exception from Moderate and Moderate Strict to Strict and  from Moderate to Moderate Strict, the stricter containment levels appear to be associated with higher numbers of new COVID-19 cases. This makes sense, albeit it is not the results we expected, when cases are rising in a country and the country's current restrictions are apparently not slowing the spread of COVID-19, it makes sense that that the country would seek to impose stricter restrictions. Only three values appear to support our hypothesis that stricter government restrictions are associated with decreasing COVID-19 cases. Those values are from Mod_Strict-Moderate (-6.819334e-06), Strict-Moderate (-1.426136e-05), and Strict-Mod_Strict (-7.442029). However, because these are the only values in support, our hypothesis cannot be confirmed.  

*** 

# Sensitivity Test

```{r,figures-side, fig.show="hold", out.width="50%"}
plot(two_way_anova)
```

## Residuals verse Fitted Plot

Based on the plot, we noticed the residual values majorly fall above the 0 line. This suggest that the assumption which the relationship is linear is not reasonable. The spreading pattern had also suggested non-equal variance of the error terms and existing outliers. 
Thus, the plot seems to indicate a non-linear model and dependency between the residuals and the fitted values. The absolute value of the residuals is strongly positively correlated with the fitted values. The data points formatted in parallel lines due to the containment index levels are categorical variables. 

## Normal Q-Q Plot

The Q-Q plot indicates the residuals deviate from the diagonal line mostly in the upper tail as the points form a steeper shape pattern. We concluded that the tail is heavier and contains larger values than what suggests under the standard modeling assumption. 
Since the residuals appear to be larger in magnitude than expected from the normal distribution, it tells us that the model might fail to comprehensively incorporate the full variability of the given data. The p-value and CI are too optimistic.  

## Scale-Location Plot

The plot shows that the residuals are not spread equally. The red smooth line is slightly steep and could be caused by the wide spread of point 45187. Scale-Location plot checks if the model violate the assumption of equal variance. Since we didn't see equally spreading points around the red smooth line shows steep trend. 

## Residuals verse Leverage Plot

This plot suggests one point with a high residual and high leverage. It appears to lie within the 0.5 and 1 line and should be further investigated. Point #45187 is identified as the influential observation. It impact the model fit and might not be approximated well by the model. If we exclude the #45187 case from our model and observe the slope coefficient and the R^2 values, we can tell how big of an impact the point actually have on our model. For now, we expect a big impact by #45187. There is no points are considered high influence as Cook's distance capture none points. 

*** 

# Discussion 

Overall, in this project we examined a fundamental question regarding the spread of COVID-19: how a country's COVID response may associate with its spread. In order to investigate this question further, we bean to create and assess various models and visuals to get a more holistic analysis of this. In our descriptive analysis portion we examined COVID-19 data from the WHO website and Oxford University on COVID-19's and drew several descriptive statistics which include number of cases, number of deaths, and spread rate. We drew further analysis by plotting histograms that showed the distribution of COVID-19 containment rates, all on a scale of 0 to 100 with 100 being maximum containment. With this we were able to calculate a quartile range to separate our data into separate containers to conduct inferential analysis. With the inferential analysis we used a two way ANOVA model and conducted a Tukey's test and found that we could not say that there were any statistically significant differences with 95% confidence. 

Having identified these findings, I believe that there need to be more studies to be able to answer this question as preventing a spread is still quite difficult for most of the world. COVID-19 has taken many lives and while with the vaccinations that are currently being given, we're on the path to preventing it, we still need to do our best to make sure we figure out the best way to prevent the spread.


***

# Acknowledgements

This report was put together with the members of Project Group 11, and facets of it were discussed with other students publicly in Piazza. Specifically, we had another user help us with an issue we were having merging our data frames.

It was constructed with the knowledge we have gathered throughout this course: various discussion sections, homework assignments, and class lectures. All instances of the lectures, especially those relating directly to the subject of this report, were of special use. The instruction and guideline for Project 1 was a good foundation point, although we deviated from that structure in minor ways.

# References

Thomas Hale, Sam Webster, Anna Petherick, Toby Phillips, and Beatriz Kira (2020). Oxford COVID-19 Government Response Tracker, Blavatnik School of Government. Retrieved March 04, 2021, https://covidtracker.bsg.ox.ac.uk/

World Bank, World Development Indicators. Available online: https://data.worldbank.org/indicator/SP.POP.TOTL (last cited: [March 4th 2021])

World Health Organization Coronavirus (COVID-19) Dashboard Geneva: World Health Organization, 2020. Available online: https://covid19.who.int/ (last cited: [March 4th 2021])




